# Introduction

This tutorial provides a step-by-step guide to running the **PEcAn Modular Workflow**. The PEcAn (Predictive Ecosystem Analyzer) system automates ecological modeling, helping researchers analyze plant functional traits, run model simulations, and perform sensitivity analyses.

After setting up PEcAn locally, open **RStudio** in your browser and log in with:
- **Username**: `carya`
- **Password**: `illinois`

Then, navigate to `web/workflow.Rmd` and open it to run the workflow. This method eliminates the need to execute the workflow through the web interface once `pecan.xml` is configured.

---

# Prerequisites

Before running this workflow, ensure you have:

- Installed **PEcAn** and its dependencies.
- An XML settings file (`settings.xml`) configured for your use case.
- A model binary (e.g., **SIPNET** or **ED2**) specified in your settings.

---

# Load PEcAn Packages

We first load the necessary PEcAn modules required for running the workflow.

```{r libraries, message=FALSE, warning=FALSE}
library("PEcAn.all")
```

---

# Load PEcAn Settings

We read the PEcAn settings from an XML file. This file contains configuration details such as input datasets, model selection, and output directories.To create a pecan.xml, you can download one generated in the PEcAn web interface or one of the `pecan.<modelname>.xml` files in the tests/ directory of 
the PEcAn repository (github.com/pecanproject/pecan).

```{r load-settings}
settings_path <- "settings.xml"  # Path to the settings file
settings <- PEcAn.settings::read.settings(settings_path)
```

---

# Prepare and Validate Settings

PEcAn provides utilities to process and validate settings before execution. This step ensures that required fields are properly set.

```{r prepare-settings}
settings$outdir <- "/home/carya/pecan_runs"
settings <- PEcAn.settings::prepare.settings(settings)
```

---

# Convert Settings to Required Units

PEcAn requires that settings be converted into the correct units before running model simulations.

```{r convert-settings}
if (!is.list(settings$host) || length(settings$host) > 1) {
  settings$host <- list(name = as.character(settings$host)[1])  # Ensure single-element list
}
settings <- PEcAn.workflow::do_conversions(settings)
```

---

# Trait and Meta Analysis

PEcAn retrieves plant trait data and performs meta-analysis to derive parameter distributions for the model.

```{r meta-analysis}
settings <- PEcAn.workflow::runModule.get.trait.data(settings)
PEcAn.workflow::runModule.run.meta.analysis(settings)
PEcAn.settings::write.settings(settings, outputfile = "pecan.TRAIT.xml")
```

---

# Write Model Configuration Files

Model-specific configuration files are generated before running simulations.

```{r run.write.configs}
settings$model$binary <- "~/pecan/models/sipnet/"  # Update the path to your model
settings <- PEcAn.workflow::runModule.run.write.configs(settings)
PEcAn.settings::write.settings(settings, outputfile = "pecan.CONFIGS.xml")
```

---

# Run Model Simulations

We execute the model using PEcAn’s workflow module.

```{r run-model}
if (!is.list(settings$host) || length(settings$host) > 1) {
  settings$host <- list(name = as.character(settings$host)[1])
}
PEcAn.utils::status.start("MODEL")
PEcAn.workflow::runModule_start_model_runs(settings)
```

---

# Model Analyses

Once the model runs are complete, we can analyze the results through sensitivity analysis and ensemble analysis.

```{r model-analysis}
runModule.get.results(settings)
run.sensitivity.analysis(settings)      # Sensitivity analysis
run.ensemble.analysis(settings)         # Ensemble analysis
run.ensemble.analysis(plot.timeseries=TRUE)  # Plot results
```

---

# Conclusion

This tutorial guided you through setting up and executing a PEcAn-based workflow. The outputs include model-generated data, sensitivity analysis results, and ensemble forecasts.


