---
title: "Uncertainty Analysis Using PEcAn"
author: "Aritra Dey"
format:
  pdf:
    toc: true
    number-sections: false
    fig-width: 10
    fig-height: 6
    fig-dpi: 300
---

# 1. Introduction {#introduction}

Welcome to this PEcAn workflow notebook! This notebook will guide you through running an ecosystem model using PEcAn's programmatic interface.

## 1.1 What Is PEcAn?

PEcAn (Predictive Ecosystem Analyzer) is a scientific workflow system designed to make ecosystem modeling more transparent, repeatable, and accessible. It helps researchers:

- Run ecosystem models with standardized inputs and outputs
- Perform uncertainty analysis on model parameters
- Compare model predictions with observations
- Share and reproduce scientific workflows

## 1.2 What This Notebook Does

This notebook demonstrates how to:

1. Set up and configure a PEcAn workflow
2. Run an ecosystem model simulation
3. Perform ensemble-based analysis to quantify uncertainty in model outputs
4. Conduct sensitivity analysis to identify which parameters most influence model results
5. Analyze and visualize the results

### The Scenario Being Modeled:

We are modeling carbon and productivity dynamics at the Niwot Ridge Forest AmeriFlux site ([US-NR1](https://ameriflux.lbl.gov/sites/siteinfo/US-NR1), a high-elevation temperate coniferous forest in Colorado. The model configuration uses the SIPNET process-based ecosystem model, parameterized with a temperate coniferous plant functional type (PFT).

The simulation is run for the full year 2004 (January 1 – December 31) using AmeriFlux LBL meteorological drivers from the Niwot Ridge site. The ensemble setup specifies one model run focusing on net primary productivity (NPP) as the target output variable.

This scenario is designed to be a minimal, reproducible example to demonstrate how to run SIPNET within the PEcAn workflow. In later steps, this same framework can be extended to include more ensemble members, additional PFTs, longer time periods, or alternative meteorological inputs.

This run is based on a study by [Moore et. al. (2007)](https://doi.org/10.1016/j.agrformet.2008.04.013) that uses SIPNET to understand the relationship between water and carbon balance at this site. 


## 1.3 Ensemble and Sensitivity Analysis in PEcAn

PEcAn can use information about parameter uncertainty to perform automated analyses:

- **Ensemble Analysis**: Repeats the model run many times, each with parameters sampled from their uncertainty distributions, to generate a probability distribution of model projections. This allows you to quantify confidence intervals on model outputs.
- **Sensitivity Analysis**: Systematically varies model parameters to assess how changes in each parameter affect model outputs. This identifies which parameters the model is most sensitive to.
- **Uncertainty Analysis**: Combines sensitivity and parameter uncertainty to determine the contribution of each parameter to the overall uncertainty in model outputs, helping to target parameters for future constraint.

These analyses are essential for understanding model behavior, quantifying uncertainty, and prioritizing data collection or parameter refinement.

## 1.4 Prerequisites

Before running this notebook, make sure you have:

- All the PEcAn packages installed. You can install all PEcAn packages and their dependencies by running the following command in the root of your PEcAn repository:

```
# Enable repository from pecanproject
options(repos = c(
  pecanproject = 'https://pecanproject.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
# Download and install PEcAn.all in R
install.packages('PEcAn.all')
```

Before running this notebook, make sure you have:

- A valid `pecan.xml` configuration file or use the example provided: `pecan/documentation/tutorials/Demo_02_Uncertainty_Analysis/pecan.xml`

## 1.5 How to Use This Notebook

1. Each section is clearly marked with a heading
2. Code chunks are provided with explanations
3. You can run the code chunks sequentially
4. Once you have successfully run the demo, you can modify parameters to configure new runs and analyses

**Objective:**

This demo illustrates how to run an uncertainty analysis workflow in PEcAn using an R-based Quarto notebook. The uncertainty workflow includes ensemble analysis and sensitivity analysis to comprehensively assess model uncertainty and parameter importance. It will cover loading settings, writing model configuration files, running model simulations, performing ensemble-based uncertainty quantification, and conducting sensitivity analysis to identify which parameters most influence model results. This approach provides a programmatic alternative to the web-based PEcAn interface for executing ecosystem models with advanced uncertainty analysis capabilities.

# 2. PEcAn Version

This section displays the version information for all PEcAn packages.

```{r version-info}
PEcAn.all::pecan_version()
```

# 3. Install SIPNET v1.3.0

If you haven't already installed the SIPNET binary, you can do so by running the following code. This will download the SIPNET binary to `demo_outdir/sipnet` and make it executable.

> Note: The `demo_outdir` directory will be created in the root of your PEcAn installation (i.e., at `documentation/tutorials/Demo_2_Uncertainty_Analysis/demo_outdir/`). This directory will contain the SIPNET binary as well as the output generated by PEcAn in this demo.

```{r download-sipnet}
# Download and install SIPNET v1.3.0
source(
  here::here(
    "documentation/tutorials/Demo_1_Basic_Run/download_sipnet.R"
  )
)
```

> Note: You can find the most recent version of the SIPNET binary at: [SIPNET GitHub Releases](https://github.com/PecanProject/sipnet/releases), but this notebook is designed to work with SIPNET v1.3.0.

# 4. Load PEcAn Packages

First, we need to load the PEcAn R packages. These packages provide all the functions we'll use to run the workflow.

```{r libraries}
# Load the PEcAn.all package, which includes all necessary PEcAn functionality
library("PEcAn.all")
```

# 5. Load PEcAn Settings Files

PEcAn uses an XML-based settings file (pecan.xml) to configure model runs. This file defines key information about the run including: PFT(s), site location, time period of the run, the location of input files and where outputs will be saved. Other settings outside the scope of this demo include the types of analyses that will be performed, how to connect to a database, and how to run it on a high performance computing cluster  (we are using the default single model run on a single computer). 

You can read more about the settings file in the ["PEcAn XML" chapter](https://pecanproject.github.io/pecan-documentation/develop/pecanXML.html#pecanXML) of the documentation.

There is an example `pecan.xml` that has been configured for this demonstration. You can find it at `pecan/documentation/tutorials/Demo_2_Uncertainty_Analysis/pecan.xml`.


This is how PEcAn loads the settings file:

```{r load-settings}
settings_path <- here::here("documentation/tutorials/Demo_02_Uncertainty_Analysis/pecan.xml")
```

# 6. Prepare and Validate Settings

After specifying the path to your `pecan.xml` file, the next step involves reading and preparing these settings. PEcAn provides utilities to process and validate your configurations before any execution begins. 

- `PEcAn.settings::read.settings(settings_path)`: Parses the `pecan.xml` file, converting its contents into an R list object that PEcAn can work with.
- `PEcAn.settings::prepare.settings(settings)`: Further prepares and validates the settings, checking for missing required fields and ensuring consistency.

```{r read-prepare-settings}
# Read the settings from the pecan.xml file
settings <- PEcAn.settings::read.settings(settings_path)

# Prepare and validate the settings
settings <- PEcAn.settings::prepare.settings(settings)
```

# 7. Explore the Settings Object

Once the settings have been read and prepared, it is useful to inspect the structure of the `settings` object. This object is an R list containing all parameters and configurations for the PEcAn workflow.

*   `str(settings)` displays the internal structure of the `settings` object. This shows how the settings are represented in R and is useful for debugging and verifying settings.

```{r explore-settings}
str(settings)
```

Your turn: explore the `settings` object further using the following commands:
*   `names(settings)` to list the top-level keys in the settings object
*   Once you know the names, you can look at each component in detail, for example:
*   `settings$run` to access the run-specific settings, such as start and end dates, model type, and output directory
*   `settings$pfts` to explore the Plant Functional Types settings

Now you can update each of these settings. Here is a simple example:

```{r add-info-to-settings}
settings$info <- list(
  author = "Aritra Dey",
  date = Sys.Date(),
  description = "Demo run of PEcAn using SIPNET"
)
```

Editing the more interesting settings to change the PFT (`settings$pfts`) or extend the run (`settings$run$end.date`) is beyond the scope of this demo. You _could_ change the pft or the end date, but you would need a new file containing parameters for that PFT (`settings$pfts$pft$posterior.files`), or a climate file (`settings$run$met$path$path1`) that extends to the desired simulation period.

# 8. Create Output Directory

Before running the workflow, we need to ensure that the output directory specified in the settings exists. This directory will store all the model outputs, configuration files, and results from the simulation.

> **Note:** All outputs will be written to `settings$outdir`. You can change this directory in your `pecan.xml` file or by modifying `settings$outdir` in the notebook before running the workflow. You can also put the `sipnet` executable anywhere you prefer, as long as you update `settings$model$binary`.

```{r create-outdir}
dir.create(settings$outdir, recursive = TRUE, showWarnings = FALSE)
```


# 9. Write Model Configuration Files

This step generates the model-specific configuration files that will be used to run the ecosystem model. The process involves:

1. Disabling database write operations because we are not using a database
2. Generating SIPNET configuration files using the `runModule.run.write.configs()` function.

```{r write-configs}
settings$database <- NULL # Disable database operations for this demo
settings <- PEcAn.workflow::runModule.run.write.configs(settings)
```

# 10. Run Model Simulations and Fetch Results

This section executes the actual model simulations and retrieves the results. The process is managed by PEcAn's workflow system, which handles the execution of your chosen ecosystem model.

*   `runModule_start_model_runs(settings)`: This function initiates the model runs based on your configuration. It manages the execution of your chosen ecosystem model, using the configuration files generated in the previous step.

```{r run-model}
PEcAn.workflow::runModule_start_model_runs(settings)
```

# 10. Ensemble and Sensitivity Analysis

This section executes ensemble and sensitivity analyses to comprehensively assess model uncertainty and parameter importance.

**Ensemble Analysis**: Quantifies uncertainty in model predictions by running multiple simulations with parameters sampled from their uncertainty distributions. This generates probability distributions of model outputs and confidence intervals.

**Sensitivity Analysis**: Systematically varies individual parameters to assess their influence on model outputs. This identifies which parameters most strongly affect model predictions and helps prioritize parameter refinement efforts.

```{r run-sensitivity-analysis}
# Execute ensemble and sensitivity analyses to quantify uncertainty and parameter influence
# First, retrieve results from all model runs
runModule.get.results(settings)

# Run ensemble analysis: generates ensemble predictions and uncertainty quantification
runModule.run.ensemble.analysis(settings)

# Run sensitivity analysis: generates variance decomposition plots and sensitivity metrics
runModule.run.sensitivity.analysis(settings)
```

# 11. Visualize Uncertainty Analysis Results

This section loads the results from the uncertainty analyses and generates plots directly in the notebook. This provides an immediate view of the ensemble time series, sensitivity plots, and variance decomposition.

```{r visualize-uncertainty-results, results='hold'}
# Load the ensemble time series data and generate the plot
ensemble_ts_file <- file.path(settings$outdir, "ensemble.ts.Rdata")
if (file.exists(ensemble_ts_file)) {
  load(ensemble_ts_file)
  PEcAn.uncertainty::plot.ensemble.ts(ensemble.ts)
}

# Load the sensitivity analysis data and generate the plots
sa_file <- file.path(settings$outdir, "sensitivity.analysis.Rdata")
if (file.exists(sa_file)) {
  load(sa_file)
  PEcAn.uncertainty::plot.sensitivity.analysis(sa.samples)
}

# Load the variance decomposition data and generate the plot
vd_file <- file.path(settings$outdir, "variance.decomposition.Rdata")
if (file.exists(vd_file)) {
  load(vd_file)
  PEcAn.uncertainty::plot.variance.decomposition(variance.decomposition.output)
}
```

# 12. Understanding PEcAn Uncertainty Analysis Outputs

After running ensemble and sensitivity analyses, PEcAn generates several important outputs that help you understand model uncertainty and parameter sensitivity.

## 12.1 Ensemble Analysis Outputs

The ensemble analysis produces:

- **`ensemble.Rdata`**: Contains `ensemble.output` object with model predictions for all ensemble members
- **`ensemble.analysis.[RunID].[Variable].[StartYear].[EndYear].pdf`**: Histogram and boxplot of ensemble predictions
- **`ensemble.ts.[RunID].[Variable].[StartYear].[EndYear].pdf`**: Time-series plot showing ensemble mean, median, and 95% confidence intervals

## 12.2 Sensitivity Analysis Outputs

The sensitivity analysis generates:

- **`sensitivity.analysis.[RunID].[Variable].[StartYear].[EndYear].pdf`**: Raw data points from univariate analyses with spline fits
- **`sensitivity.output.[RunID].[Variable].[StartYear].[EndYear].Rdata`**: Model outputs corresponding to parameter variations

## 12.3 Uncertainty Analysis Outputs

The uncertainty analysis produces:

- **`variance.decomposition.[RunID].[Variable].[StartYear].[EndYear].pdf`**: Three-column analysis showing:
  - Coefficient of variation (normalized posterior variance)
  - Elasticity (normalized sensitivity) 
  - Partial standard deviation of each parameter

## 12.4 Interpreting the Results

**Variance Decomposition Analysis:**
- Parameters are sorted by their contribution to model output variability
- Identify top-tier parameters for future constraint
- Distinguish between parameters that are:
  - Highly sensitive but low uncertainty
  - Highly uncertain but low sensitivity  
  - Both sensitive and uncertain (priority targets)

**Parameter Prioritization:**
- Focus on parameters that are both sensitive and uncertain
- Consider measurement costs and feasibility
- Balance parameter uncertainty vs. model sensitivity

# 13. Customizing Ensemble Analysis Parameters (Optional)

## (Optional) Use this section only if you want to override the default ensemble analysis parameters. Skip if defaults are sufficient.
> **Important:** If you modify the ensemble analysis parameters in this section, you must re-run sections 8, 9, and 10 to regenerate the model output with the new ensemble configuration.

```{r customize-ensemble}
# Set the number of ensemble members (model runs)
settings$ensemble$size <- 50

# Specify the variable to be analyzed in the ensemble
settings$ensemble$variable <- "NEE"

# Choose the method for sampling the parameter space (options: "uniform", "lhc", "halton", "sobol", "torus")
settings$ensemble$samplingspace$parameters$method <- "halton"
```


# 14. Customizing Sensitivity Analysis Parameters (Optional)

## 

