---
title: "PEcAn Modular Workflow"
author: "Aritra Dey"
date: "`r Sys.Date()`"
output: html_document
---

# Load PEcAn packages
```{r libraries}
library("PEcAn.all")
```
# Load PEcAn settings files.

Open and read in settings file for PEcAn run.
To create a pecan.xml, you can download one generated in the PEcAn web interface or one of the `pecan.<modelname>.xml` files in the tests/ directory of the PEcAn repository (github.com/pecanproject/pecan).
```{r load-settings}

settings_path <- "settings.xml" 
settings <- PEcAn.settings::read.settings(settings_path)
```


```{r prepare-settings}
# Prepare and check settings
settings <- PEcAn.settings::prepare.settings(settings)

if ("benchmarking" %in% names(settings)) {
  library(PEcAn.benchmark)
  settings <- papply(settings, read_settings_BRR)
}

# Process sitegroup settings
if ("sitegroup" %in% names(settings)) {
  if (is.null(settings$sitegroup$nSite)) {
    settings <- PEcAn.settings::createSitegroupMultiSettings(settings,
      sitegroupId = settings$sitegroup$id
    )
  } else {
    settings <- PEcAn.settings::createSitegroupMultiSettings(
      settings,
      sitegroupId = settings$sitegroup$id,
      nSite = settings$sitegroup$nSite
    )
  }
}
```

```{r convert-settings}
# Convert settings to required units
if (!is.list(settings$host)) {
  settings$host <- list(name = settings$host)
}
settings <- PEcAn.workflow::do_conversions(settings)

if (is.null(settings$debug)) {
  settings$debug <- FALSE  # Default to FALSE if not defined
}
```

# Trait and Meta Analysis
Fetches plant trait data and prior distributions from external databases for the specified Plant Functional Types (PFTs), then combines trait data across multiple sources to derive probabilistic distributions for model parameters.
If this step is skipped, a posterior file should be specified in the settings.

```{r meta-analysis}
if (settings$debug) cat("Starting trait and meta-analysis...\n")
# Retrieve trait data and prior distributions for the specified Plant Functional Types (PFTs)
settings <- PEcAn.workflow::runModule.get.trait.data(settings)
# Perform meta-analysis to derive probabilistic distributions for model parameters
PEcAn.workflow::runModule.run.meta.analysis(settings)
# Save the updated settings, including the retrieved trait data, to an XML file
PEcAn.settings::write.settings(settings, outputfile = "pecan.TRAIT.xml")

```



# Write Model Configuration Files
Generates model-specific configuration files based on the settings, ensuring proper model execution.

```{r run.write.configs}
settings <- runModule.run.write.configs(settings)
```


# Run Model and Retrieve Results
Executes the model simulations and retrieves the results for further analysis.

```{r run-model}
if (!is.list(settings$host)) {
  settings$host <- list(name = settings$host)
}
PEcAn.utils::status.start("MODEL")
runModule.start.model.runs(settings)
```

# Model Analyses
Performs various analyses on model outputs, including sensitivity analysis, variance decomposition, and ensemble analysis.

```{r model-analysis}
runModule.get.results(settings)
run.sensitivity.analysis()      # Run sensitivity analysis and variance decomposition on model output
run.ensemble.analysis()  	      # Run ensemble analysis on model output. 
run.ensemble.analysis(plot.timeseries=TRUE) 
```

