---
title: "PEcAn Modular Workflow"
author: "Aritra Dey"
date: "`r Sys.Date()`"
output: html_document
---

# Load PEcAn packages
```{r libraries}
library("PEcAn.all")
```
# Load PEcAn settings files.

Open and read in settings file for PEcAn run.
To create a pecan.xml, you can download one generated in the PEcAn web interface or one of the `pecan.<modelname>.xml` files in the tests/ directory of the PEcAn repository (github.com/pecanproject/pecan).
```{r load-settings}

settings_path <- "settings.xml" 
settings <- PEcAn.settings::read.settings(settings_path)
settings <- PEcAn.utils::do_conversions(settings)

if (is.null(settings$debug)) {
  settings$debug <- FALSE  # Default to FALSE if not defined
}
```

# Trait Analysis
Fetches plant trait data and prior distributions from external databases for the specified Plant Functional Types (PFTs)

```{r trait-analysis}
if (settings$debug) cat("Starting trait analysis...\n")
# Retrieve trait data and prior distributions for the specified Plant Functional Types (PFTs)
settings <- PEcAn.workflow::runModule.get.trait.data(settings)
# Save the updated settings, including the retrieved trait data, to an XML file
PEcAn.settings::write.settings(settings, outputfile = "pecan.TRAIT.xml")

```

# Meta Analysis
Combines trait data across multiple sources to derive probabilistic distributions for model parameters

```{r meta-analysis}
if (settings$debug) cat("Starting meta-analysis...\n")
# Perform meta-analysis to derive probabilistic distributions for model parameters
settings <- runModule.get.trait.data(settings)
runModule.run.meta.analysis(settings)
```

# Write Model Configuration Files
Generates model-specific configuration files based on the settings, ensuring proper model execution.

```{r run.write.configs}
settings <- runModule.run.write.configs(settings)
```


# Run Model and Retrieve Results
Executes the model simulations and retrieves the results for further analysis.

```{r run-model}
runModule.start.model.runs(settings)
runModule.get.results(settings)
```

# Model Analyses
Performs various analyses on model outputs, including sensitivity analysis, variance decomposition, and ensemble analysis.

```{r model-analysis}
run.sensitivity.analysis()      # Run sensitivity analysis and variance decomposition on model output
run.ensemble.analysis()  	      # Run ensemble analysis on model output. 
run.ensemble.analysis(plot.timeseries=TRUE) 
```


```{r finish}
if (settings$debug) {
    cat("Workflow completed successfully\n")
}
```
